---
title: "Lab 3"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
library(shiny)
library(shinythemes)
library(shinydashboard)
library(plotly)
library(emojifont)
library(dplyr)
library(shinyWidgets)
```
In this lab you need to create a shiny app. 

Naturally, you would need to install `shiny` package.  You can either create ui.R and server.R files, or just one .R file with both UI and server codes in
it.

```{r}
ibm_data <- read.csv2("WA_Fn-UseC_-HR-Employee-Attrition.csv", sep= ",")

#Data categories


ibmGender <- select(ibm_data, HourlyRate, Gender, )
ibmDepartment <- select(ibm_data, HourlyRate, Department)
ibmAge <- select(ibm_data, HourlyRate, Age)



ibmAtt <- character()
for(i in 1:length(ibm_data$Attrition)){
  if (ibm_data$Attrition[i] == "Yes"){
    ibmAtt[i] <- 'Yes'
  }else if(ibm_data$Attrition[i] == "No"){
    ibmAtt[i] <-'No'
  }
}


ibmAttAge <- na.omit(cbind(ibmAge,ibmAtt))
ibmAttDepartment<- na.omit( cbind(ibmDepartment,ibmAtt))
ibmAttGender <- na.omit(cbind(ibmGender,ibmAtt))

m = mean(ibmAge$HourlyRate)
s = sd(ibmAttAge$HourlyRate)



```

## Requirements

- (10 points) The app should be about some statistics topic or technique of your choice (for example: regression with various data sets, simulation of various probability distributions, confidence and prediction bands in regression, etc.).

- (10 points) At least two different kinds of GUI components/widgets must be included (slider, checkbox, table, tabs, menubar/selectbar, radio button, submit button, etc.)

- (20 points) Must include an interactive plot (you can use base plot, ggplot, or (3D) plotly)

- (10 points) Shiny app must knit for grader (so make sure to turn in all components the app needs to run), must be turned in on time, must include PDF of HTML output and RMD file.  You also need to submit in Canvas a PDF file explaining what your shiny app does, what its features are, etc. The description of your app doesn’t have to be detailed, especially if it is clear from the plot. If, however, your app has some interesting feature that is not obviously visible upon opening the app, you may want to point that out and explain it.

## Note

Apart from examples of shiny apps in Canvas, you can also go to shiny gallery (https://shiny.rstudio.com/gallery/) and see a lot of cool apps, or you can run the following examples in you RStudio console.
```{r}
ui<- fluidPage(
  dashboardBody(setBackgroundImage(src ="https://images.pexels.com/photos/735911/pexels-photo-735911.jpeg?auto=compress&cs=tinysrgb&w=1260&h=750&dpr=2" )),
  theme = shinytheme("superhero"),
  navbarPage("IBM Attrition Rate Analysis", 
             tabPanel("Introduction", 
                      mainPanel(tags$h1("Introduction",emoji('book')),
                                p("\"Employee attrition is  defined as the natural process by which employees leave the workforce.\"(BasuMallick, 2021). IBM has created a dataset to analyze employee attrition rates vs performance.  "),
                                p("This app will use normal distribution to display the correlation between employee attrition, age, gender, and department vs hourly rate."),
                                p("All data used in this application can be found here: https://www.kaggle.com/datasets/pavansubhasht/ibm-hr-analytics-attrition-dataset"),
                                
                                tags$h3("Features", emoji('rocket'),emoji('computer')),
                                p('This app features the following:'),
                                p("A Navagational Bar."),
                                p("An interactive Plotly bar plot."),
                                p("A custom royalty free stock photo background."),
                                p("Radio button selection for various graphical comparisons."),
                                
                                tags$h2("References"),
                                p("BasuMallick. (2021, March 11). What Is Employee Attrition? Definition, Attrition Rate, Factors, and Reduction Best Practices. Https://Www.Pexels.Com/Photo/Photo-of-Person-Typing-on-Computer-Keyboard-735911/. https://www.toolbox.com/hr/engagement-retention/articles/what-is-attrition-complete-guide/Photo of Person Typing on Computer Keyboard. (2017, October 15). [Photograph]. https://images.pexels.com/photos/735911/pexels-photo-735911.jpeg?auto=compress&cs=tinysrgb&w=1260&h=750&dpr=2")
                                
               
             )),
             tabPanel("Normal Distribution",
                      sidebarPanel(tags$h3("Attrition Filter"),
                       
                          radioButtons("att_choice",
                                       label= "Filter",
                                       choiceNames = list("Age","Gender","Department"),
                                       choiceValues = list('Age','Gender','Department')
                            
                          )
                        
                      ),
                      
                      mainPanel(
                        plotlyOutput("plot")
                        )  
                      ),
             )

   )


```

```{r}
server<- function(input,output){
  
observeEvent(input$att_choice,{
  output$plot <- switch(input$att_choice,
                        'Age' = renderPlotly({plot_ly(ibmAttAge, 
                                                      x =ibmAttAge$HourlyRate,
                                                      type="histogram", 
                                                      color = ibmAttAge$ibmAtt)%>%
                                                      layout(title = 'Normal Distribution', 
                                                             xaxis = list(title = 'Hourly Rate'), 
                                                             yaxis = list(title = 'Employee Age'), 
                                                             legend = list(title=list(text='<b> Attrition </b>')))%>%
                                                              add_trace(x = rnorm(ibmAttAge$HourlyRate, mean = m, sd = s))}),
                        'Gender' =  renderPlotly({plot_ly(ibmAttGender, 
                                                          x = ibmAttGender$Gender,
                                                          y= ibmAttGender$HourlyRate,
                                                          color = ibmAttGender$ibmAtt,
                                                          type="bar")}),
                        
                        'Department' = renderPlotly({plot_ly(ibmAttDepartment, 
                                                             x = ibmAttDepartment$Department, 
                                                             y= ibmAttDepartment$HourlyRate, 
                                                             color = ibmAttDepartment$ibmAtt,
                                                             type="bar")}))
  output$att_choice <-renderText(input$att_choice)})
}



shinyApp(ui = ui,server = server)

```

